Indexing Variation Graphs\thanks{This work was supported by the Wellcome Trust grant 098051.}

### 2016-08-10 paper/paper.tex
Started preparing a new version of the paper

\documentclass[a4paper,UKenglish]{lipics-v2016}

\usepackage{microtype}

\bibliographystyle{plainurl}

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Indexing Variation Graphs\footnote{This work was supported by the Wellcome Trust grant 098051.}}

\author[1]{Jouni Sirén}
\affil[1]{Wellcome Trust Sanger Institute, Hinxton, Cambridge, UK\\
  \texttt{jouni.siren@iki.fi}}
\authorrunning{J. Sirén}

\Copyright{Jouni Sirén}

\subjclass{E.1 Data Structures}
\keywords{Burrows-Wheeler transform, de Bruijn graphs, path indexes, variation graphs}
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\EventEditors{John Q. Open and Joan R. Acces}
%\EventNoEds{2}
%\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
%\EventShortTitle{CVIT 2016}
%\EventAcronym{CVIT}
%\EventYear{2016}
%\EventDate{December 24--27, 2016}
%\EventLocation{Little Whinging, United Kingdom}
%\EventLogo{}
%\SeriesVolume{42}
\ArticleNo{1}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% DNA
\newcommand{\dnaseq}[1]{\ensuremath{\mathtt{#1}}}
\newcommand{\baseA}{\dnaseq{A}}
\newcommand{\baseC}{\dnaseq{C}}
\newcommand{\baseG}{\dnaseq{G}}
\newcommand{\baseT}{\dnaseq{T}}
\newcommand{\baseN}{\dnaseq{N}}
\newcommand{\dnacomp}[1]{\ensuremath{\overline{#1}}}
\newcommand{\revcomp}[1]{\ensuremath{\overleftarrow{#1}}}

@@

@@

\maketitle

@@

\section{Introduction}

@@

We primarily consider sequences over the \emph{DNA} alphabet $\set{\$, \baseA, \baseC, \baseG, \baseT, \baseN}$. Characters $\baseA$, $\baseC$, $\baseG$, and $\baseT$ are called \emph{bases}, while character $\baseN$ represents an arbitrary or unknown base. In addition to the endmarker $\$$, the alphabet may also contain other characters for technical purposes. Each character $c$ of the DNA alphabet has a \emph{complement} $\dnacomp{c}$ defined as $\dnacomp{\baseA} = \baseT$, $\dnacomp{\baseC} = \baseG$, $\dnacomp{\baseG} = \baseC$, $\dnacomp{\baseT} = \baseA$, and $\dnacomp{c} = c$ for other characters $c$. Given a DNA sequence $S$, its \emph{reverse complement} is the sequence $\revcomp{S}$ obtained by reversing the non-technical parts of the sequence and replacing each character with its complement. For example, $\revcomp{\dnaseq{GATTACA}\$} = \dnaseq{TGTAATC}\$$.

@@

\begin{figure}[t!]
\includegraphics{gcsa2_text_indexes.pdf}
\caption{\LFmapping, BWT, suffix array, and LCP array for text $\dnaseq{GCATCATA}\$$.}\label{figure:indexes}
\end{figure}

@@

\begin{figure}[t!]
\includegraphics[width=396pt]{gcsa2_graph_dbg.pdf}
\caption{Left: Input graph $G = (V, E)$, with each node $v \in V$ labeled with $v:G.\glabel(v)$. Right: The \orderk{3} de~Bruijn graph $G' = (V', E')$ of graph $G$, with each node $v' \in V'$ labeled with $G'.\gkey(v')$ and $G'.\gvalue(v')$. Both: Edges $(t, s)$ are not shown. The highlighted path in the de~Bruijn graph is a false positive, as it consists of two disjoint paths in the input graph.}\label{figure:graph-dbg}
\end{figure}

@@

\begin{figure}[t!]
\includegraphics[width=396pt]{gcsa2_pruned_index.pdf}
\caption{Left: An \orderk{3} pruned de~Bruijn graph $G''$ \kequivalent{3} to the de~Bruijn graph in Figure~\ref{figure:graph-dbg}. Right: GCSA for graph $G''$. Leftward arrows show how backward searching proceeds, with the red arrows demonstrating it from pattern $\dnaseq{T}$ to pattern $\dnaseq{AT}$. Rightward arrows tell where the samples for each node are located, with the blue ones showing them for node $\dnaseq{CAT}$.}\label{figure:pruned-index}
\end{figure}

@@

\subparagraph*{Acknowledgements.}

@@

\bibliography{paper}

>>>>

\documentclass[a4paper,11pt]{llncs}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{fullpage}
\setlength\tabcolsep{4pt}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{textgreek}
\usepackage{microtype}
\usepackage{hyperref}

@@

% DNA
\newcommand{\baseA}{\mathtt{A}}
\newcommand{\baseC}{\mathtt{C}}
\newcommand{\baseG}{\mathtt{G}}
\newcommand{\baseT}{\mathtt{T}}
\newcommand{\baseN}{\mathtt{N}}
\newcommand{\dnacomp}[1]{\ensuremath{\overline{#1}}}
\newcommand{\revcomp}[1]{\ensuremath{\overleftarrow{#1}}}

@@

\title{Indexing Variation Graphs\thanks{This work was supported by the Wellcome Trust grant 098051.}}

\author{Jouni Sirén}
\institute{Wellcome Trust Sanger Institute, UK \\ \email{jouni.siren@iki.fi}}

@@

\thispagestyle{empty}
\setcounter{page}{0}
\maketitle

@@

\clearpage
\section{Introduction}

@@

We primarily consider sequences over the \emph{DNA} alphabet $\set{\$, \baseA, \baseC, \baseG, \baseT, \baseN}$. Characters $\baseA$, $\baseC$, $\baseG$, and $\baseT$ are called \emph{bases}, while character $\baseN$ represents an arbitrary or unknown base. In addition to the endmarker $\$$, the alphabet may also contain other characters for technical purposes. Each character $c$ of the DNA alphabet has a \emph{complement} $\dnacomp{c}$ defined as $\dnacomp{\baseA} = \baseT$, $\dnacomp{\baseC} = \baseG$, $\dnacomp{\baseG} = \baseC$, $\dnacomp{\baseT} = \baseA$, and $\dnacomp{c} = c$ for other characters $c$. Given a DNA sequence $S$, its \emph{reverse complement} is the sequence $\revcomp{S}$ obtained by reversing the non-technical parts of the sequence and replacing each character with its complement. For example, $\revcomp{\mathtt{GATTACA}\$} = \mathtt{TGTAATC}\$$.

@@

\begin{figure}[t!]
\includegraphics{gcsa2_text_indexes.pdf}
\caption{\protect\LFmapping, BWT, suffix array, and LCP array for text $\mathtt{GCATCATA}\$$.}\label{figure:indexes}
\end{figure}

@@

\begin{figure}[t!]
\includegraphics[width=396pt]{gcsa2_graph_dbg.pdf}
\caption{Left: Input graph $G = (V, E)$, with each node $v \in V$ labeled with $v:G.\glabel(v)$. Right: The \protect\orderk{3} de~Bruijn graph $G' = (V', E')$ of graph $G$, with each node $v' \in V'$ labeled with $G'.\gkey(v')$ and $G'.\gvalue(v')$. Both: Edges $(t, s)$ are not shown. The highlighted path in the de~Bruijn graph is a false positive, as it consists of two disjoint paths in the input graph.}\label{figure:graph-dbg}
\end{figure}

@@

\begin{figure}[t!]
\includegraphics[width=396pt]{gcsa2_pruned_index.pdf}
\caption{Left: An \protect\orderk{3} pruned de~Bruijn graph $G''$ \protect\kequivalent{3} to the de~Bruijn graph in Figure~\ref{figure:graph-dbg}. Right: GCSA for graph $G''$. Leftward arrows show how backward searching proceeds, with the red arrows demonstrating it from pattern $\mathtt{T}$ to pattern $\mathtt{AT}$. Rightward arrows tell where the samples for each node are located, with the blue ones showing them for node $\mathtt{CAT}$.}\label{figure:pruned-index}
\end{figure}

@@

\paragraph*{Acknowledgements.}

@@

\bibliographystyle{plainurl}
\bibliography{paper}
